treats manual

### 2023-11-22 inst/gitbook/04_make.events.Rmd
change time.condition to age.condition
For bypassing s3 issues on the CRAN

function | arguments | input | what does it do
---------|-----------|-------|----------------
`time.condition`, `taxa.condition`, `trait.condition` | x | the value to check the target against | e.g. time = x, taxa = x or trait = x (see below)
 | condition | a relational operator | e.g. `==`, `>`, `!=`, etc. for asking for example `time == x` 
 | ... | any additional arguments to be passed to the event | |

@@

 * `time.condition` will trigger the `event` once a certain time is reached in the simulations and is the simplest/most basic condition with no arguments other than the time required. For example ``time.condition(4, condition = `==`)`` will trigger the `event` once the simulations reaches 4 time units. Easy.
 * `taxa.condition` will trigger the `event` once a certain number of taxa are reached. This can be considered including or excluding fossil species. For example ``taxa.condition(42, condition = `>=`, living = TRUE)`` will trigger the `event` if there are at least 42 living taxa.
 * `trait.condition` will trigger the `event` once a certain trait value is reached. This function allows to say which trait(s) to target (by default, the first one using `trait = 1`), what value of the trait to target (by default `what = max`) and whether to use and absolute trait value or not (`absolute = TRUE`). For example ``trait.condition(1/3, condition = `>`, trait = 1, what = sd)`` will trigger the condition after the standard deviation of the first trait reaches 1/3.

@@

```{r}
## 80% mass extinction at time 4
mass_extinction <- make.events(
                      target = "taxa",
                      condition = time.condition(4),
                      modification = random.extinction(0.8))

>>>>

function | arguments | input | what does it do
---------|-----------|-------|----------------
`age.condition`, `taxa.condition`, `trait.condition` | x | the value to check the target against | e.g. time = x, taxa = x or trait = x (see below)
 | condition | a relational operator | e.g. `==`, `>`, `!=`, etc. for asking for example `time == x` 
 | ... | any additional arguments to be passed to the event | |

@@

 * `age.condition` will trigger the `event` once a certain time is reached in the simulations and is the simplest/most basic condition with no arguments other than the time required. For example ``age.condition(4, condition = `==`)`` will trigger the `event` once the simulations reaches 4 time units. Easy.
 * `taxa.condition` will trigger the `event` once a certain number of taxa are reached. This can be considered including or excluding fossil species. For example ``taxa.condition(42, condition = `>=`, living = TRUE)`` will trigger the `event` if there are at least 42 living taxa.
 * `trait.condition` will trigger the `event` once a certain trait value is reached. This function allows to say which trait(s) to target (by default, the first one using `trait = 1`), what value of the trait to target (by default `what = max`) and whether to use and absolute trait value or not (`absolute = TRUE`). For example ``trait.condition(1/3, condition = `>`, trait = 1, what = sd)`` will trigger the condition after the standard deviation of the first trait reaches 1/3.

@@

```{r}
## 80% mass extinction at time 4
mass_extinction <- make.events(
                      target = "taxa",
                      condition = age.condition(4),
                      modification = random.extinction(0.8))

### 2023-11-22 inst/gitbook/07_specific_examples.Rmd
updated manuals with returns
change time.condition to age.condition
For bypassing s3 issues on the CRAN

Functionality showcase | Example
--------------|----------
`time.condition` | [mass extinction](#EGrandom_ext), [negative trait extinction](#EGneg_ext), [reducing speciation rate](#EGred_spec), [changing trait process](#EG_change_trait), [changing modifiers](#EG_change_modif)
`random.extinction` | [mass extinction](#EGrandom_ext)
`trait.extinction` | [negative trait extinction](#EGneg_ext)
`taxa.condition` | [background extinction](#EGbg_ext), [change branch length](#EG_modify_brlen)
`bd.params.update` | [background extinction](#EGbg_ext), [reducing speciation rate](#EGred_spec)
`traits.update` | [changing trait process](#EG_change_trait), [changing trait correlation](#EG_change_correlation)
`trait.condition` | [changing trait correlation](#EG_change_correlation), [generating a subtree with a different process](EG_founding_traits)
`modifiers.update` | [changing modifiers](#EG_change_modif), [change branch length](#EG_modify_brlen)
`parent.traits` | [changing modifiers](#EG_change_modif)
`founding.event` | [generating a subtree with no extinction](#EG_founding_purebirth), [generating a subtree with a different process](EG_founding_traits)

@@

We will then create a event that triggers when reaching half of the simulation time (using `time.condition(2.5)`).
This event will target `"taxa"`, i.e. the number of species, and randomly remove 80% of them (using `random.extinction(0.8)`):

```{r}
## 80% mass extinction at time 4
mass.extinction <- make.events(
                    condition = time.condition(4),
                    target = "taxa",
                    modification = random.extinction(0.8))
```

@@

And our extinction event which triggers after reaching time 4 (`time.condition(4)`), targets the `"taxa"` and modifies the extinction for species with traits lower than 0.

```{r}
## Extinction of any tips with trait < 1 at time 4
trait.extinction <- make.events(
                      target = "taxa",
                      condition = time.condition(4),
                      modification = trait.extinction(x = 0,
                                                      condition = `<`))
```

@@

And our extinction event which triggers after reaching time 4 (`time.condition(4)`), targets the `"bd.params"` and modifies the speciation rate to 1/3:

```{r}
## Reducing speciation after reaching time 4
reduced.speciation <- make.events(
                      condition = time.condition(4),
                      target = "bd.params",
                      modification = bd.params.update(speciation = 1/3))
```

@@

And our event which triggers after reaching time 5 (`time.condition(5)`), targets the `"traits"` and modifies the process to `OU.process`.

```{r}
## Create an event to change the trait process
change.process.to.OU <- make.events(
                  condition    = time.condition(5),
                  target       = "traits",
                  modification = traits.update(process = OU.process))
```

@@

## Event for changing a modifier: speciation event increase for species with negative values {#EG_change_modif}

@@

And our extinction event which triggers after reaching time 3 (`time.condition(3)`), targets the `"modifiers"` and modifies birth-death rule as follows:
 * When a species is descendant from a parent with a negative trait value (`negative.trait.condition`);
 * Then increase your chances of going extinct by +1 (`increase.extinction.1`)

@@

## Update the modifier
change.speciation <- make.events(
    condition    = time.condition(3),
    target       = "modifiers",
    modification = modifiers.update(speciation = speciation,
                                    condition  = negative.trait.condition,
                                    modify     = increase.extinction.1))
```

>>>>

Functionality showcase | Example
--------------|----------
`age.condition` | [mass extinction](#EGrandom_ext), [negative trait extinction](#EGneg_ext), [reducing speciation rate](#EGred_spec), [changing trait process](#EG_change_trait), [changing modifiers](#EG_change_modif)
`random.extinction` | [mass extinction](#EGrandom_ext)
`trait.extinction` | [negative trait extinction](#EGneg_ext)
`taxa.condition` | [background extinction](#EGbg_ext), [change branch length](#EG_modify_brlen)
`bd.params.update` | [background extinction](#EGbg_ext), [reducing speciation rate](#EGred_spec)
`traits.update` | [changing trait process](#EG_change_trait), [changing trait correlation](#EG_change_correlation)
`trait.condition` | [changing trait correlation](#EG_change_correlation), [generating a subtree with a different process](EG_founding_traits)
`modifiers.update` | [changing modifiers](#EG_change_modif), [change branch length](#EG_modify_brlen)
`parent.traits` | [changing modifiers](#EG_change_modif)
`founding.event` | [generating a subtree with no extinction](#EG_founding_purebirth), [generating a subtree with a different process](EG_founding_traits)

@@

We will then create a event that triggers when reaching half of the simulation time (using `age.condition(2.5)`).
This event will target `"taxa"`, i.e. the number of species, and randomly remove 80% of them (using `random.extinction(0.8)`):

```{r}
## 80% mass extinction at time 4
mass.extinction <- make.events(
                    condition = age.condition(4),
                    target = "taxa",
                    modification = random.extinction(0.8))
```

@@

And our extinction event which triggers after reaching time 4 (`age.condition(4)`), targets the `"taxa"` and modifies the extinction for species with traits lower than 0.

```{r}
## Extinction of any tips with trait < 1 at time 4
trait.extinction <- make.events(
                      target = "taxa",
                      condition = age.condition(4),
                      modification = trait.extinction(x = 0,
                                                      condition = `<`))
```

@@

And our extinction event which triggers after reaching time 4 (`age.condition(4)`), targets the `"bd.params"` and modifies the speciation rate to 1/3:

```{r}
## Reducing speciation after reaching time 4
reduced.speciation <- make.events(
                      condition = age.condition(4),
                      target = "bd.params",
                      modification = bd.params.update(speciation = 1/3))
```

@@

And our event which triggers after reaching time 5 (`age.condition(5)`), targets the `"traits"` and modifies the process to `OU.process`.

```{r}
## Create an event to change the trait process
change.process.to.OU <- make.events(
                  condition    = age.condition(5),
                  target       = "traits",
                  modification = traits.update(process = OU.process))
```

@@

## Event for changing a modifier: extinction event increase for species with negative values {#EG_change_modif}

@@

And our extinction event which triggers after reaching time 3 (`age.condition(3)`), targets the `"modifiers"` and modifies birth-death rule as follows:
 * When a species is descendant from a parent with a negative trait value (`negative.trait.condition`);
 * Then increase your chances of going extinct by +1 (`increase.extinction.1`)

@@

## Update the modifier
change.speciation <- make.events(
    condition    = age.condition(3),
    target       = "modifiers",
    modification = modifiers.update(speciation = speciation,
                                    condition  = negative.trait.condition,
                                    modify     = increase.extinction.1))
```
